---
title: "Lab 4 Team Collaboration"
author: "Andrew Maloney & Elaine Ob"
date: "3/28/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading Data
```{r}

library(readxl)
library(dplyr) 
library(forcats)
library(ggplot2)
library(tidyr)
library(stringr)
library(reshape2)
defense <- read_excel('cyclonesFootball2018.xlsx', sheet = 'Defensive')

offense  <- read_excel('cyclonesFootball2018.xlsx', sheet = 'Offensive')

bio <- read_excel('cyclonesFootball2018.xlsx', sheet = 'Biography')


```

## Part One: Cleaning data

#### 1. The player names and opponent names are factors.
#### 2. All offensive and defensive statistics are numerical. The Weight column in biogrpahy is also numeric.
#### 3. Bonus: Change the Height column in biogrpahy into numeric.
```{r}
#str(defense)

defClean <- defense %>% mutate(Name = factor(Name), Opponent_Opponent = factor(Opponent_Opponent)) %>% mutate_at(vars(starts_with("Tack")), list(as.numeric)) %>% mutate_at(vars(starts_with("Turn")), list(as.numeric)) %>% mutate_at(vars(starts_with("Pass")), list(as.numeric))

str(defClean)
print(defClean)

```

```{r}
#str(offense)

offClean <- offense %>% mutate(Name = factor(Name), Opponent_Opponent = factor(Opponent_Opponent)) %>% mutate_at(vars(starts_with("Rec")), list(as.numeric)) %>% mutate_at(vars(starts_with("Rush")), list(as.numeric)) %>% mutate_at(vars(starts_with("Pass")), list(as.numeric))

str(offClean)
print(offClean)
```

```{r}

bioClean <- bio %>% mutate(Weight = as.numeric(Weight)) %>% separate(Height, c("Feet", "Inches"), "-", convert = TRUE) %>% mutate(Height_in_Centimeters = ((Feet*12) + Inches)*2.54)

str(bioClean)
print(bioClean)
```


## Part Two: Tidying(1)
#### 1. Reformat the defClean data frame into a tidy format using gather. The type of statistic (Tackles_Solo, Tackles_ASST, etc) is added as a new key column named Statistic.
```{r}

defClean_Tidy <- defClean %>% gather(key = Statistic, value = Count, 3:11)
print(defClean_Tidy)
```

#### 2. Create a histogram for each defensive statistic. Use facet_wrap to draw all histograms in a single plot. What defensive skills are rare?
```{r}
defClean_Tidy %>% ggplot(aes(x = Count, fill = Statistic))+ 
  geom_histogram(binwidth = 0.7) + facet_wrap(~Statistic)+
  ggtitle("Defense Statistic histogram") + xlab("Defense Statistic")

```

Looking at the 9 histograms we can see that the Turnovers are a rare defensive skill.  This makes sense to anyone who has watched a football game.  Turnover plays do not happen often.

#### 3. Did ISU have better defense against West Virginia or Kansas State?
```{r}
Teams <- c("West Virginia", "Kansas State")
defClean_Tidy_Teams <- defClean_Tidy %>% filter(Opponent_Opponent %in% Teams) %>% spread(key = Opponent_Opponent, value = Count) %>% filter(Statistic == "Tackles_Solo")

print(defClean_Tidy_Teams)

defClean_Tidy_Teams %>% ggplot(aes(x = `West Virginia`, y = `Kansas State`))+ 
  geom_point(na.rm = TRUE) + geom_jitter(na.rm = TRUE)+ 
  ggtitle("Which Game did Iowa State play better Defense?") + xlab("Solo Tackles againt West Virginia") +ylab("Solo Tackles againt Kansas State")

#defClean %>% select(Name, Opponent_Opponent, Tackles_Solo) %>% spread(Opponent_Opponent, Tackles_Solo) %>% ggplot(aes(x = `West Virginia`, y = `Kansas State`)) + geom_point()
```

Looking at the newly created data frame and scatterplot, we can see Iowa State had a significanlty better defense game against Kansas State then when playing West Virginia.  

## Part Two: Tidying(2)
#### 1. Separate the Hometown column into two columns, namely the city and the state. Print the first few rows of these two columns. 

#### 2. How many players are there from each state?
```{r}
#str(bioClean)

bioClean_Tidy <- bioClean %>% separate(Hometown, c("City", "State"), ",", convert = TRUE )
print(bioClean_Tidy, 3)

Players_State <- table(bioClean_Tidy$State)
print(Players_State)

```



## Part Three: Joing Dataframes 
#### 1. Find where do the offensive players come from by joining the offensive data and player biography data appropriately.
```{r}
Joined_Offense_with_Biography <- inner_join(offClean, bioClean_Tidy, by = "Name")
Players_Origin <- Joined_Offense_with_Biography %>% distinct(Name, City, State, Highschool)
print(Joined_Offense_with_Biography)
print(Players_Origin)


```

#### 2. Is there a relationship between the weight of a player and the receiving yards (Receiving_YDS)?
```{r}
Joined_Offense_with_Biography %>% ggplot(aes(x = Weight, y = Receiving_YDS, color = Name))+
  geom_point(na.rm = TRUE) + geom_jitter(na.rm = TRUE)

```

There does not seem to be a relationship between the weight of a player and the amount of receiving yards.  It is more likely that there is a relationship between the position the player plays and the amount of receiving yards.

#### 3. Explore more. Raise a question that can be answered only by looking at two or more data frames. Then answer that question.

New Question: Is there anything we can conclude from the Height and Weight of a player to the positions of a player on Iowa State Football Team? 
```{r}
Joined_Defense_with_Biography <- inner_join(defClean, bioClean_Tidy, by = "Name")

Complete_Join_Biography <- full_join(Joined_Defense_with_Biography, Joined_Offense_with_Biography)
Complete_Join_Biography_2 <- Complete_Join_Biography %>% distinct(Name, Position, Weight, Height_in_Centimeters)
print(Complete_Join_Biography_2)

Complete_Join_Biography_2 %>% ggplot(aes(x = Position, y = Height_in_Centimeters, fill = Position))+
  geom_boxplot(na.rm = TRUE)+
  ggtitle("Height Box Plot") + xlab("Position") +ylab("Height in Centimeters") + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
  
Complete_Join_Biography_2 %>% ggplot(aes(x = Position, y = Weight, fill = Position))+
  geom_boxplot(na.rm = TRUE)+
  ggtitle("Weight Box Plot") + xlab("Position") +ylab("Weight in Pounds") + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
```

Cornerbacks tend to be smaller players with a smaller stature and lower weight. Deep snappers seems to be taller and normal weight. Defensive backs tend to be smaller players with a lower weight and height. Defensive ends have a higher weight and taller than normal. Defensive line players have a normal height with a higher weight. Line backers seem to have a height around 6'0-6'2" and a weight of 210 to 235 pounds.

Offensive line players seems to be very tall and weigh the most out of all the positions. Place kicker tend to have a low weight with an average height. QB's tend to be weigh less but are taller. Running backs tend to weigh less (but more than the defensive backs) and are shorter than other positions. Tight ends are tall and have a higher weight. Wide receivers tend to weigh less but are very tall.